---
title: "kmdata - intro"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{kmdata - intro}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, include=FALSE}
library('knitr')

opts_chunk$set(
  cache = FALSE, fig.align = 'center', dev = 'png',
  fig.width=9, fig.height=7, echo = TRUE, message = FALSE
)
```

## Data sets available

```{r}
library('kmdata')

data(package = 'kmdata')
```

This will show a list of the data sets available. Data objects can be
references by `ATTENTION_2A` or `` `TRIBE(2)_2A` `` (with backticks) for data
names with special characters.

The name consists of the study short-name and figure identifier:

```{r}
data <- ls('package:kmdata', pattern = '^[A-Z]')
cbind(
  name = data,
  study = gsub('_.*', '', data),
  figure = gsub('.*_', '', data)
)[1:5, ]
```

For example, study "ATTENTION" has two figures: "2A" and "2B." Study "ACTSCC"
has only one figure: "2A." If data sets sourced from multi-panel figures, the
name will look similar to study "ATTENTION" with figure IDs "2A" and "2B."

All data sets are listed in `kmdata_key` along with some useful metadata for
each including the journal and publication identifiers, outcomes and study
arms, quality of the re-capitulated data, and other information.

## Working with the data

Each data set contains the same format for consistency:

```{r, echo=FALSE}
knitr::kable(
  data.frame(
    time = 'time-to-event (in units)', event = 'event indicator (0/1)',
    arm = 'treatment arm identifier (e.g., arm-1 vs arm-2)'
  )
)
```

The time unit, event type, and treatment arms can be found in the help page
for each data set, e.g., `?ACT1_2A`. Additionally, the data objects contain
metadata stored as attributes:

```{r}
head(ACT1_2A)

attr(ACT1_2A, 'event')

attributes(ACT1_2A)[-(1:3)]
```

Data may be examined and plotted using the built-in functions `summary` and
`kmplot`.

```{r, fig.show='hold', message = TRUE}
summary(ACT1_2A)

kmplot(ACT1_2A)
```

## Selecting data

The `kmdata` package contains a function, `select_kmdata`, to easily search
and filter data sets which share common features. Any of the columns in
`kmdata_key` may be used to filter.

For example, if we wanted a list of lung cancer data sets with overall
survival (OS) in months with fewer than 500 patients reporting at least a
1.2 hazard ratio for treatment compared to a reference arm, we can use the
following:

```{r}
select_kmdata(
  Cancer %in% 'Lung' &
    Outcome %in% 'OS' &
    Units %in% 'months' &
    ReportedSampleSize < 500 &
    HazardRatio >= 1.2,
  return = 'name'
)
```

By default, `select_kmdata` returns only the names of the data sets for
reference individually (i.e., `select_kmdata(..., return = 'name')`), but
it can also return the matching rows of `kmdata_key` or the matching data
sets as a list.

```{r, fig.height=11, fig.width=12, results='hide'}
key <- select_kmdata(
  Cancer %in% 'Lung' &
    Outcome %in% 'OS' &
    Units %in% 'months' &
    ReportedSampleSize < 500 &
    HazardRatio >= 1.2,
  return = 'key'
)

dat <- select_kmdata(
  Cancer %in% 'Lung' &
    Outcome %in% 'OS' &
    Units %in% 'months' &
    ReportedSampleSize < 500 &
    HazardRatio >= 1.2,
  return = 'data'
)

par(mfrow = n2mfrow(length(dat)))
for (dd in dat)
  kmplot(dd)
```

## Data quality

Each figure and data set contains a quality score which represents how well
the re-capitulated agrees with the original publication. Scores range from
0 (worst) to 100% (best) and are an aggregation of four metrics: hazard ratio,
total events, median time-to-event, and number at-risk.

Each metric is score from 0 (worst) to 3 (best); the maximum score per figure
may vary with the metrics reported in the original publication. For example,
if only one was reported, the maximum score is 3/3.

A score of 3 points is given per metric per figure if the re-capitulated
metric is no more than 5% different than the published, 2 points are given
if the metric is 5-10% different, 1 point for 10-20%, and 0 points for more
than 20% different.

|  % difference from publication |  Quality points per metric |
|-------------------------------:|---------------------------:|
|                            0-5 |                          3 |
|                           5-10 |                          2 |
|                          10-20 |                          1 |
|                        &gt; 20 |                          0 |

---

## References

The publications and figures available in this package are listed below by
first author.

<details>
<summary>Click to expand</summary>
```{r, echo=FALSE}
cit <- system.file('docs', 'Citations_final.xlsx', package = 'kmdata')
cit <- as.data.frame(readxl::read_excel(cit, skip = 1L))

cit <- within(cit, {
  Title    <- gsub('^.*?\\.\\s+|\\.\\s+[A-z ]+\\d{4};.*$', '', Reference)
  Author   <- gsub('^([^.]+\\.)|.', '\\1', Reference)
  PubData  <- gsub('([A-z ]+\\s+\\d{4};.*)\\.$|.', '\\1', Reference)
  Journal  <- gsub('(.*?)\\d{4};|.', '\\1', PubData)
  Year     <- gsub('(\\d{4});|.', '\\1', PubData)
  Location <- gsub('^.*?\\d{4};\\s*', '\\1', PubData)
})[, c('PMID', 'Author', 'Journal', 'Year', 'Title', 'Location')]
cit <- cit[order(cit$Author), ]
rownames(cit) <- NULL

knitr::kable(cit, format = 'markdown', caption = 'List of publications.')
```
</details>
