---
title: "kmdata - ipd"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{kmdata - ipd}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, include=FALSE}
library('knitr')

opts_chunk$set(
  cache = FALSE, fig.align = 'center', dev = 'png',
  fig.width=9, fig.height=7, echo = TRUE, message = FALSE
)
```

## Generating IPD

In addition to the recapitulated study data, the `kmdata` package contains
the algorithm to re-create patient-level time-to-event data from digitized
survival curves.

Digitized curves are simply the x- and y-coordinates of the Kaplan-Meier
curves which can be generated several ways including the \code{R} package
**`digitize`** or proprietary software such as
[DigitizeIt](https://digitizeit.soft112.com/).

```{r}
library('kmdata')

xy <- system.file('etc', 'init', 'Checkmate 067_S3A_Nivolumab.csv',
                  package = 'kmdata')
ar <- system.file('etc', 'init', 'At Risk.csv',
                  package = 'kmdata')

dd <- read.csv(xy)
aa <- read.csv(ar)
aa <- aa[aa$nrisk > 0, ]

ipd <- ipd(dd$T, dd$S, aa$trisk, aa$nrisk, arm = 'Nivo')
kmplot(ipd, conf.int = FALSE, median = FALSE)

points(S ~ T, dd, type = 's', col = 'red')

legend(
  'topright', legend = c('Re-capitulated', 'Truth (no censoring\ndata available)'),
  col = 1:2, lty = 1, bty = 'n', lwd = 2
)
```

## References

Guyot, Patricia, et al. Enhanced secondary analysis of survival data:
reconstructing the data from published Kaplan-Meier survival curves.
_BMC Medical Research Methodology_ 2012, **12**:9.